EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-05-14 draft.tex
Title proposal following @jj recommendation
consistent docker-compose name

  \title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a container-based application that reactively processes
isolated and
heterogeneous populations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

@@

Although at first sight containers look like simply a method that
allows easy shipping of applications, they are much more than
that. One of the first features that make them interesting is their
ephemerality: fast startup and teardown time make them usable for
machine-independent and cloud deployment of simple, ephemeral and
maybe stateless functions. On the other hand, containerization also
allows, through the use of orchestration engines like Docker Swarm or
Kubernetes, or through simple, non-self-scaling specifications using
Docker Compose, a definition of all the relationships between
different parts of the application, or services. The full realization
of Docker-based architectures has led to the creation of a series of
patterns that, together, conform cloud-native application development
\cite{kratzke2017understanding}.

@@

% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
% (and he next one too) - JJ
% done - Mario
% But also change title? - JJ
% Experimental design of a container-based application or something like that - JJ
\section{Container-based application} 
\label{docker}

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. % several issues that make it practical or useful instead? - JJ
In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same system could be
deployed to the cloud.

% The main advantage of container deployments over virtual machine
% instances is that they add much less overhead to communication between
% them, and they are much more economical to deploy. Scaling and
% monitoring can be done much more easily through cloud orchestration
% systems such as Kubernetes; these are the reasons why we chose it for
% this work.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ
% OK, I will add an issue - Mario
% Did you finally? - JJrehe
% I did not, will add and assign it - M
There are many advantages of container-based architectures over one
based on virtual machines. The main ones are architectural: it makes
the design of an application gravitate towards a set of decoupled
tasks, that use reactive, event based programming, to interact with
the rest of the application; containers hold micro-services that are
much easier to test, integrate and deploy. It also simplifies
the interaction APIs. Of course, that makes them more economical in a
pay-per-use cloud provider, but that is not the main reason why we
have chosen them in this paper; reproducibility of experiments and a
clear and idempotent description of the architecture through docker
compose % Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ
are the main reasons.

@@

\begin{table}[h!tbp]
  \small
  \caption{ Experiment configuration example 
  }
  \label{tab:params}
  \centering
  \small
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
    \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
    \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
    \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
    \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
    \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
    \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
    \multicolumn{3}{|l|}{Mutation  Probability Range}  & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
    \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
    Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
            10              & 50      & 140                 &      5                 & 30                \\ \hline
            20              & 66      & 200                 &      5                 & 30               \\ \hline
            40              & 80      & 250                 &      5                 & 40                \\ \hline
  \end{tabular}
\end{table}

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the docker-compose file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ

@@

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run docker-compose commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster.

@@

  \item[A container orchestration service] these services are compatible with docker-compose
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.

@@

As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration 
(from deploying it locally using docker-compose) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ

@@

As we can see, the local environment (see section \ref{setup}) achieved a higher
ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single
docker-compose file, and one command to execute. But with the limit of 4CPUs and
ten containers, it is the slower option. The second option (labeled Multi), is
faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
containers, it is reasonable to think that we can achieve a higher number of
evaluations with an appropriate configuration. Nevertheless, the goal of this
experiment was not to increase the number of evaluations but to test the
flexibility of a cloud-native application.

@@

% this is not a conclusion. The conclusion is an answer to the
% research question. This works, and works precisely for the reasons
% we say it works. - JJ
% Again, revise search questions and explain answers - JJ
As a conclusion, EvoSwarm is a framework designed and tested according
to industry standards which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.

>>>>

  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
  % I like cloud-native more than event-based. 
  % How about that one?  - M

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a container-based application that reactively processes
isolated and
heterogeneous populations. The platform uses the {\tt docker-compose} tool for
defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

@@

Although at first sight containers look like simply a method that
allows easy shipping of applications, they are much more than
that. One of the first features that make them interesting is their
ephemerality: fast startup and teardown time make them usable for
machine-independent and cloud deployment of simple, ephemeral and
maybe stateless functions. On the other hand, containerization also
allows, through the use of orchestration engines like Docker Swarm or
Kubernetes, or through simple, non-self-scaling specifications using
{\tt docker-compose}, a definition of all the relationships between
different parts of the application, or services. The full realization
of Docker-based architectures has led to the creation of a series of
patterns that, together, conform cloud-native application development
\cite{kratzke2017understanding}.

@@

% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
% (and he next one too) - JJ
% done - Mario
% But also change title? - JJ
% Experimental design of a container-based application or something like that - JJ 
% Like it - M
\section{Experimental design of a container-based application} 
\label{docker}

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience, several
issues that make this approach more practical. % several issues that make it practical or useful instead? - JJ 
In general, cloud platforms have free-tiers    
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same system could be
deployed to the cloud.

% The main advantage of container deployments over virtual machine
% instances is that they add much less overhead to communication between
% them, and they are much more economical to deploy. Scaling and
% monitoring can be done much more easily through cloud orchestration
% systems such as Kubernetes; these are the reasons why we chose it for
% this work.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ
% OK, I will add an issue - Mario
% Did you finally? - JJrehe
% I did not, will add and assign it - M
There are many advantages of container-based architectures over one
based on virtual machines. The main ones are architectural: it makes
the design of an application gravitate towards a set of decoupled
tasks, that use reactive, event based programming, to interact with
the rest of the application; containers hold micro-services that are
much easier to test, integrate and deploy. It also simplifies
the interaction APIs. Of course, that makes them more economical in a
pay-per-use cloud provider, but that is not the main reason why we
have chosen them in this paper; reproducibility of experiments and a
clear and idempotent description of the architecture through {\tt docker-compose} 
% Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ - ok
are the main reasons.

@@

\begin{table}[h!tbp]
  \small
  \caption{ Experiment configuration example 
  }
  \label{tab:params}
  \centering
  \small
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
    \multicolumn{3}{|l|}{Worker Containers}        & \texttt{int}     & \texttt{8} \\ \hline
    \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
    \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
    \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
    \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
    \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
    \multicolumn{3}{|l|}{Mutation  Probability Range}  & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
    \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
    Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
            10              & 50      & 140                 &      5                 & 30                \\ \hline
            20              & 66      & 200                 &      5                 & 30               \\ \hline
            40              & 80      & 250                 &      5                 & 40                \\ \hline
  \end{tabular}
\end{table}

@@

For local experiments we have deployed the EvoSwarm application in a
Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ

@@

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run {\tt docker-compose} commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster.

@@

  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.

@@

As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration 
(from deploying it locally using {\tt docker-compose}) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ

@@

As we can see, the local environment (see section \ref{setup}) achieved a higher
ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
network overhead in a local workstation and a clock speed of 3.8 GHz. The first
option (labeled Single) is straightforward to deploy, needing only a single
{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
ten containers, it is the slower option. The second option (labeled Multi), is
faster, and not too far from the local execution, especially when solving the
$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
containers, it is reasonable to think that we can achieve a higher number of
evaluations with an appropriate configuration. Nevertheless, the goal of this
experiment was not to increase the number of evaluations but to test the
flexibility of a cloud-native application.

@@

% this is not a conclusion. The conclusion is an answer to the
% research question. This works, and works precisely for the reasons
% we say it works. - JJ
% Again, revise search questions and explain answers - JJ
As a conclusion, EvoSwarm is a framework designed and tested according
to industry standards which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and {\tt docker-compose} installed.
